The Globular Cluster System of the Auriga Simulations

### 2019-03-26 AurigaGCS.tex
Some fixes for data section, and split what we see (results) from what we interprete (discussion)

\section{Relevant observational data}
\label{sec:observations}
We summarise relevant observations of the globular cluster system of the Milky 
Way in Sec.~\ref{sec:milkyway}, and of Andromeda (M31) in Sec.~\ref{sec:andromeda}.

@@

\subsubsection*{Age estimates}
We supplement H96e10 with age-estimates from isochrone fits to stars
near the main-sequence turnoff in 55 GCs \citep[][hereafter V13]{2013ApJ...775..134V}.
The mean value of the age-estimates in this data set is $11.9 \pm 0.1$~Gyr and 
the dispersion is $0.8$~Gyr. Furthermore, only one of the 55 GC age-estimates is
below $10$~Gyr.

\subsubsection*{Distribution of total GC mass in metallicity-radial bins}
In Figure~\ref{fig:MW-FeHRgc} we show the two-dimensional mass-weighted 
metallicity-radial distribution of the MW GCS. In Sec.~\ref{sec:FeHRgc} 
we investigate whether the star formation model implemented in the Auriga 
simulations can produce sufficient total mass in GC candidates in the same bins.

\begin{figure}
    \includegraphics[width=\columnwidth]
        {{MW_RgcFeH_HistogramMassWeighted_Harris1996ed2010data}.png}
    \includegraphics[width=\columnwidth]
        {{M31_RgcFeH_HistogramMassWeighted_CaldwellRomanowsky2016data}.png}
    \caption{
        \emph{Top}: Mass-weighted R$_{\text{GC}}$-[Fe/H] distribution of
        151 GCs in the MW \citep[data from][2010 ed.]{1996AJ....112.1487H}, which
        is 98.19 \% of the total MW GCS mass in the Harris catalog.
        \emph{Bottom}: Same for M31, showing 366 GCs and 88.38 \% of the total 
        mass in CR16 \citep[data from][]{2016ApJ...824...42C}. Note that the
        range of the colourmap differs in both figures.
        \label{fig:MW-FeHRgc}
    }
\end{figure}

@@

C11 and CR16 present a uniform set of spectroscopic observations calibrated 
on the Milky Way GCS of the inner $1.6^\circ~(\sim21)$~kpc that 
is believed to be 94\% complete. GCs in the outer stellar halo, up to 
$R_{\text{proj}}\sim150$~kpc, are observed in the Pan-Andromeda Archaeological
Survey \citep[PAndAS, ][hereafter H14]{2014MNRAS.442.2165H}, but see also
\citet{2014MNRAS.442.2929V} and \citet{2019MNRAS.484.1756M}. H14 presents the
discovery of 59 new GCs and publishes updates to RBC~5. The work of H14 is
incorporated in the latest public release\footnote{Last revised 23 Sep 2015, see
\url{https://www.cfa.harvard.edu/oir/eg/m31clusters/M31_Hectospec.html}} 
of the C11 dataset, further revised by CR16. It seems that CR16 is the most 
recent aggregated dataset of M31's GCS that contains properties of interest
for our study as it contains GCs in the inner region and in the outer halo. The 
relevant fields in the CR16 dataset that we use are the age, metallicity, and the 
mass-estimate\footnote{The authors assumed $M/L_V = 2$ (independent of [Fe/H])}.

@@

\subsubsection*{Age estimates}
For M31 we find an age distribution with a mean value of 11.0$\pm 0.2$~Gyr and a 
dispersion of $2.2$~Gyr. Furthermore, $24$ GCs have age-estimates below $10$~Gyr, and
the minimum age is $4.8$~Gyr.

@@

\subsubsection*{Calculation of galactocentric radius}
The projected galacocentric radius $R_{\text{proj}}$ is calculated following
\citet[][Sec.~4.1, eq.~4]{2019arXiv190111229W}. Specifically,
\begin{align}
    X &= A \sin{(\text{PA})} + B \cos{(\text{PA})} \nonumber \quad \text{and} \\
    Y &= - A \cos{(\text{PA})} + B \sin{(\text{PA})} \text{,}
\end{align}
where $A = \sin{(\alpha - \alpha_0)} \cos{\delta}$ and
$B = \sin\delta \cos\delta_0 - \cos{(\alpha - \alpha_0)} \cos\delta \sin\delta_0$
for the right ascension $\alpha$ (hourangle) and declination $\delta$ (degree) 
of the GCs. For the central position of M31 we take the preferred position 
from the NASA Extragalactic Database\footnote{\url{https://ned.ipac.caltech.edu/}}
entry of M31: $(\alpha_0, \, \delta_0) = 
(0^{\text{h}}42^{\text{m}}44.35^{\text{s}}, \, +41^{\circ}16'08.63")$.
The projected galactocentrid radius is calculated as $R_\text{proj} = 
\sqrt{X^2 + Y^2}$ (for $X$ and $Y$ in arcsec), and using the small angle 
approximation with a distance of $D_{\text{M31}} = 778$~kpc to convert to kpc.
TODO: McConnachie et al. 2005; Conn  et  al. 2012. Finally, we convert the 
projected distance ($R_{\text{proj}}$) to an `average deprojected distance` 
via the relationship $R_{\text{GC}} = R_{\text{proj}} \times (4/\pi)$.

@@

We consider the metallicity distribution in Sec.~\ref{sec:metallicity} 
the distribution of galactocentric radii in Sec.~\ref{sec:radial}, and the
combination of both in Sec.~\ref{sec:FeHRgc}.

\subsection{Metallicity distribution}
\label{sec:metallicity}
We investigate whether the star formation model implemented in the Auriga 
simulations is capable of i) producing a metallicity distribution that is 
consistent with the GCS observed in the MW (M31), and ii) generating sufficient 
mass over the domain of the observations.

\begin{figure}
    \includegraphics[width=\columnwidth]{{FeH_mu_sigma}.png}
    \caption{
        \label{fig:FeH_mu_sigma}
    }
\end{figure}

Figure~\ref{fig:FeH} shows a mass-weighted metallicity distribution of the
Auriga L4 median
for \emph{all stars} (orange dotted), \emph{old stars / GC candidates} 
(orange solid), \emph{old in-situ} stars (blue solid), and \emph{old accreted} stars
(red solid). The shaded regions indicate the $1\sigma$ interval around the median 
(i.e. the scatter between runs with different initial conditions and merger
histories). The MW (M31) GCS is shown in purple (magenta). We use the 
same bin sizes for the simulations as for the observations, explicitly shown for
the observed profiles.

@@

\subsection{Radial distribution}
\label{sec:radial}

@@

Is the spatial distribution of the GC candidates in the Auriga simulations consistent with the MW GCS?

\subsection{Metallicity-radial distribution}
\label{sec:FeHRgc}
What age-metallicity distribution is produced by star formation events in the Auriga simulations?
?
\begin{figure}
    \includegraphics[width=0.49\textwidth]
        {{Au-all_RgcFeH_HistogramMassWeighted_iold_mean}.png}
    \caption{
        Mass-weighted [Fe/H]-R$_{\text{GC}}$ distribution of all Auriga haloes 
        (level 3, 4 and 5). Here we consider the old ($>$~10~Gyr) stars in all 
        simulations and color-code the \textbf{mean value} (of 40 Auriga haloes)
        \label{fig:Au-FeHRgc}
    }
\end{figure}

\begin{figure}
    \includegraphics[width=0.49\textwidth]
        {{Au-all_RgcFeH_HistogramMassWeighted_iold_36bins_mean-trim}.png}
    \caption{
        Ratio
        \label{fig:Au-FeHRgc-ratio}
    }
\end{figure}

\subsection{Metallicity-radial distribution: higher resolution}
\todo[inline]{Do we even want this?}
\label{sec:FeHRgc-morebins}
\begin{figure}
    \includegraphics[width=\columnwidth]
        {{logMRgcFeH_accreted-insitu-notitle}.png}
    \caption{
        TODO: add 1$\sigma$ interval around both relations, and flip x and y.
        \label{fig:todo2}
    }
\end{figure}

@@

\subsection{Properties of birth haloes of the accreted population}
\label{sec:birth-properties-of-accreted-gc-candidates}

Auriga galaxy /w 5 Myr snapshots.

\subsection{Age-metallicity distribution}
\label{sec:agemetallicity}
What age-metallicity distribution is produced by star formation events in the Auriga simulations?

\subsection{Formation history} 
\label{sec:history}
Can we identify particular star formation events that generate GC candidates with the correct age, metallicity, and radial properties as expected or the MW GCS?

Can we distinguish between particles that have formed in-situ and those that have been accreted? Can we identify specific features in the age-metallicity plane, or in the $R_\textbf{GC}$-[Fe/H] plane, that result from one of both populations? How does this connect to proposed mechanisms for globular cluster formation in the literature?

Orbits: are the pericentres different? Look at velocity + specific angular momentum distribution in the different FeH/Rgc bins as proxy for the pericenter

\section{Discussion}
\label{sec:discussion}

We investigate all star particles in the Auriga simulations that are older than 10~Gyr,
an approach equal to the method of \citep{2017MNRAS.465.3622R}. This approach does not
take the bound cluster fraction \citep[e.g.][]{2012MNRAS.426.3008K} into account. This
means that our sub set, which is based on selection by age, comprises both stars in the
field as well as globular clusters. We compare the total mass in the simulations in 
metallicity ([Fe/H]), galactocentric radius $R_{\text{GC}}$, and [Fe/H]-$R_{\text{GC}}$ 
bins to the total mass in the MW GCS (using the H96e10 data set). The mass excess in the
simulations gives a maximum mass loss `budget' for the product of cluster formation 
efficiency and dynamical evolution. We find fractions that vary with metallicity,
radius and metallicity-and-radius.

>>>>

\section{Observational data}
\label{sec:observations}
We summarise relevant observations of the globular cluster system of the Milky 
Way in Sec.~\ref{sec:milkyway}, and of Andromeda (M31) in Sec.~\ref{sec:andromeda}.

@@

\subsubsection{Age estimates}
We supplement H96e10 with age-estimates from isochrone fits to stars
near the main-sequence turnoff in 55 GCs \citep[][hereafter V13]{2013ApJ...775..134V}.
The mean value of the age-estimates in this data set is $11.9 \pm 0.1$~Gyr and 
the dispersion is $0.8$~Gyr. Furthermore, only one of the 55 GC age-estimates is
below $10$~Gyr.

@@

C11 and CR16 present a uniform set of spectroscopic observations calibrated 
on the Milky Way GCS of the inner $1.6^\circ~(\sim21)$~kpc that 
is believed to be 94\% complete. GCs in the outer stellar halo, up to 
$R_{\text{proj}}\sim150$~kpc, are observed in the Pan-Andromeda Archaeological
Survey \citep[PAndAS, ][hereafter H14]{2014MNRAS.442.2165H}, but see also
\citet{2014MNRAS.442.2929V} and \citet{2019MNRAS.484.1756M}. H14 presents the
discovery of 59 new GCs and publishes updates to RBC~5. The work of H14 is
incorporated in the latest public release\footnote{Last revised 23 Sep 2015, see
\url{https://www.cfa.harvard.edu/oir/eg/m31clusters/M31_Hectospec.html}} 
of the C11 dataset, further revised by CR16. It seems that CR16 is the most 
recent aggregated dataset of M31's GCS that contains properties of interest
for our study as it contains GCs in the inner region and in the outer halo. The 
relevant fields in the CR16 dataset that we use are the age, metallicity, and the 
mass-estimate\footnote{The authors assumed $M/L_V = 2$ independent of [Fe/H]}.

@@

\subsubsection{Age estimates}
For M31 we find an age distribution with a mean value of 11.0$\pm 0.2$~Gyr and a 
dispersion of $2.2$~Gyr. Furthermore, $24$ GCs have age-estimates below $10$~Gyr, and
the minimum age is $4.8$~Gyr.

@@

\subsection{Total GC mass in metallicity-radial space}
\label{sec:observations_FeHRgc}
We show the two-dimensional mass-weighted metallicity-radial distribution of the 
MW (M31) GCS in the top (bottom) panel of Figure~\ref{fig:observations_FeHRgc}. Both
quantities are readily available in H96e10, but the galactocentric radius
of GCs in M31 is not available in CR16. Therefore we follow 
\citet[][Sec.~4.1]{2019arXiv190111229W} to calculate the projected radius 
$R_{\text{proj}}$ from the observed positions, adopting M31's central position
from the NASA Extragalactic Database\footnote{\url{https://ned.ipac.caltech.edu/}} 
$(\alpha_0, \, \delta_0) = 
(0^{\text{h}}42^{\text{m}}44.35^{\text{s}}, \, +41^{\circ}16'08.63")$ 
and distance $D_{\text{M31}} = 778$~kpc (McConnachie et al. 2005; Conn  et  al. 2012).
We calculate $R_{\text{GC}}$ as `average deprojected distance`
$R_{\text{GC}} = R_{\text{proj}} \times (4/\pi)$.
We compare these observations to the Auriga simulations later on in 
Section~\ref{sec:discussion_FeHRgc},

\begin{figure}
    \includegraphics[width=\columnwidth]
        {{MW_RgcFeH_HistogramMassWeighted_Harris1996ed2010data}.png}
    \includegraphics[width=\columnwidth]
        {{M31_RgcFeH_HistogramMassWeighted_CaldwellRomanowsky2016data}.png}
    \caption{
        \emph{Top}: Mass-weighted R$_{\text{GC}}$-[Fe/H] distribution of
        151 GCs in the MW \citep[data from][2010 ed.]{1996AJ....112.1487H}, which
        is 98.19 \% of the total MW GCS mass in the Harris catalog.
        \emph{Bottom}: Same for M31, showing 366 GCs and 88.38 \% of the total 
        mass in CR16 \citep[data from][]{2016ApJ...824...42C}. Note that the
        range of the colourmap differs in both figures.
        \label{fig:observations_FeHRgc}
    }
\end{figure}

@@

We consider the metallicity distribution in Sec.~\ref{sec:results_FeH} 
the distribution of galactocentric radii in Sec.~\ref{sec:results_Rgc}, and the
combination of both in Sec.~\ref{sec:FeHRgc}.

\subsection{Metallicity distribution}
\label{sec:results_FeH}
We investigate whether the star formation model implemented in the Auriga 
simulations is capable of i) producing a metallicity distribution that is 
consistent with the GCS observed in the MW (M31), and ii) generating sufficient 
mass over the domain of the observations.

\begin{figure}
    \includegraphics[width=\columnwidth]{{FeH_mu_sigma}.png}
    \caption{
        Mean and standard deviation of the metallicity distribution of star
        particles in our Auriga subsets compared to the MW (M31) GCS shown in
        purple (magenta). The black solid (open) dot shows values of the 
        metal-rich (metal-poor) component of the MW GCS as quoted by
        \citep[][p. 38]{1998gcs..book.....A}.
        \label{fig:FeH_mu_sigma}
    }
\end{figure}

Figure~\ref{fig:FeH_mu_sigma} shows the mean and standard deviation of the
[Fe/H] distribution of star particles in the Auriga simulations. The mean
metallicity of old star particles lies in the range $-1.0$ to $-0.5$ and the
dispersion increases with decreasing metallicity. For comparison, the mean
of all star particles populates the range $-0.25$ to $0$ with a near-constant
dispersion a bit above $0.5$. Old accreted stars in the Auriga haloes reach
mean metallicities between $-1.5$ and $-1.0$, which is considerably lower than 
the old- and old insitu populations. Most Auriga L4 haloes have old insitu 
populations with mean metallicities around $-0.5$. The old accreted populations 
follow the trent of increasing dispersion with decreasing mean metallicity, but 
notably the old insitu populations all have a constant dispersion of roughly
$0.45$. We show the values calculated using all MW (M31) GCs in purple (magenta),
and literature values of the metal-rich (metal-poor) component of the MW GCS
as black solid (open) dots.

Figure~\ref{fig:FeH} shows the mass-weighted metallicity distribution of the
Auriga L4 haloes. Lines show the median value and the shaded regions indicate 
the $1\sigma$ interval (i.e. scatter between runs with different initial 
conditions which thus have unique merger histories). We notice once more that
the peak metallicity shifts from $0$ to $-0.5$ for old stars (orange solid)
compared to all stars (orange dotted).
(orange solid), \emph{old in-situ} stars (blue solid), and \emph{old accreted} stars
(red solid). 
of \emph{all stars} (orange dotted),

The MW (M31) GCS is shown in purple (magenta). We use the 
same bin sizes for the simulations as for the observations, explicitly shown for
the observed profiles.

@@

\subsection{Radial distribution}
\label{sec:results_Rgc}

@@

\begin{figure}
    \includegraphics[width=0.49\textwidth]
        {{Au-all_RgcFeH_HistogramMassWeighted_iold_mean}.png}
    \caption{
        Mass-weighted [Fe/H]-R$_{\text{GC}}$ distribution of all Auriga haloes 
        (level 3, 4 and 5). Here we consider the old ($>$~10~Gyr) stars in all 
        simulations and color-code the \textbf{mean value} (of 40 Auriga haloes)
        \label{fig:Au-FeHRgc}
    }
\end{figure}

\subsection{Total mass in metallicity-radial space}
\label{sec:results_FeHRgc}
\begin{figure}
    \includegraphics[width=0.49\textwidth]
        {{Au-all_RgcFeH_HistogramMassWeighted_iold_36bins_mean-trim}.png}
    \caption{
        Ratio
        \label{fig:Au-FeHRgc-ratio}
    }
\end{figure}

\section{Discussion}
\label{sec:discussion}
\subsection{Metallicity distribution}
\label{sec:discussion_FeH}
\citet[][p. 234]{1998gcs..book.....A} and \citet[][p. 38]{Harris2001} describe
the bimodal [Fe/H] distribution of the MW GCS. The latter fits a double
Gaussian that peaks at [Fe/H] = $-1.59$ (metal-poor) and $-0.51$ (metal-rich)
with dispersions of $0.34$ and $0.23$. T
KMM mixture-modeling

\subsection{Radial distribution}
\label{sec:discussion_Rgc}

\subsection{Metallicity-radial space}
\label{sec:discussion_FeHRgc}
See Figure~\ref{fig:observations_FeHRgc} in Section~\ref{sec:observations_FeHRgc}.

@@
